---
title: "data_manipulation"
author: "mggn"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

### Data manipulation
you want to select relevant variables, filter out unnecessary obs,
create new vars, change existing ones, arrange in an easy to digest format

dplyr package allows us to select, filter, mutate, and arrange
modularity is intentional, and these functions share a structure

piping allows you to tie together a sequence actions, come from the magrittr 
package

PIPES, sequence of actions
in R, you can NEST actions like do(brush(wakeup(Maria)))
like those russian nesting dolls
you can alternatively do it in a sequence
--> piping is a solution to this
%>% represents "and then"
happy = wake_up(maria) %>% brush_teeth() %>% drink_covfefe()

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
select(litters_df, -litter_number) #aka keep everything except for litter #
```

we can select what we want to keep by using syntax select(dataset, var, var)
rename and select can do similar things
select(dataset, startswith(")), similar to in SAS using contains fx
relocate and select can also do similar things

## filter
select operates on columns, filter operates on rows

filter(litters_df, gd_weight < 22) this is of course similar to.. the filter option!
== test for equality 
!= refers to not equal to
%in% 

##mutate
```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
  )
```

##arrange

```{r}
arrange(litters_df, pups_born_alive)
#ways to organize the data

```

## DA PIPE %>%

```{r}
litters_df =
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names()%>%
  select(-pups_survive) %>%
  mutate(wt_gain = gd18_weight - gd0_weight) %>%
  drop_na(gd0_weight)
```
