---
title: "data_wrangling_lecture"
author: "mggn"
date: "9/22/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl) #to import excel datasets
library(haven) #to import SAS datasets
```

There are two kinds of paths:

-Absolute: a file or folder’s “full address” on your computer
-Relative: directions to a file or folder from your current working directory

RELATIVE paths are preferable, as they are portable
this means that, if you share a relative path then the code won't "break"
Absolute paths are specific to your computer!

Let's now try to read in some data:

```{r data}
litters_df = read_csv("./data/FAS_litters.csv")
litters_def = janitor::clean_names(litters_df)

#janitor::clean_names takes the clean_names function from the janitor package
#instead of having to load the whole janitor package
```

Take a look at the data by just printing in the cosole
```{r}
litters_def
```


```{r}
tail(litters_def)
```

```{r}
skimr::skim(litters_def)
```

To view the dataset in a new window, we can use the command view() BUT this is
best used by typing into the console, and not putting it in a code chunk in the 
r markdown file. This is because including a command that opens up a new window
in an R markdown file can cause issues when knitting

write_csv(subdata, "./data/mlbsubtable.csv") <-- this EXPORTS data!