---
title: "tidy_data"
author: "mggn"
date: "9/22/2020"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Tidy data
data tables have an implied structure which the tidy data framework 
  makes explicit
columns are variables
rows are observations
every value has a cell

Benefits: consistent data structures will simplify your thought process, data
written for a computer is easier to work with
**we learned this in the SQL class!
If you keep things tidy, the data cleaning process will ben easier :)

NOT ALL DATA are tidy
-columns are values, not var names
-single columns contain multiple variables
-data are stored in multiple tables
-data written for humans is generally not tidy
-some data aren't really amenable to tidiness (genomics, neuroimaging)

RELATIONAL data
-data spread across tables with defined relations
-variables used to define these relations are keys
-tables are combined by joins (inner or outer joins)

## pivot_longer

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
```

Wide format to lng format

```{r}
pulse_tidy_data =
  pulse_data %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi"
  )

```

rewrite, combine, expand


